---
title: "Covid19 vaccine study Fall2021"
author: "Paulo Poblete"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```

## Barplot 
```{r}
barchartGC(~treatment + infected ,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")
```

As per the plot above, it shows us that those who took the placebo got more sick than the males who took the Covid91. We can surmise that the "sick" portion of the placebo group is almost double compared to the drug group. This shows that the drug is effective for the Males group. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
As per the table above, 33.11 of the sick males had the treatment, and 66.89 of those males who got sick had the placebo. There is a 2:1 difference. This lets us know that the treatment is indeed working. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

As per the chi sq test above, the p value is very tiny. This suggest that we need to reject the null hypothesis and that there is a difference between the males who took the drug in comparison to those who took the placebo. 

```{r}
MaleEstimate <-Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <-Malesmod$conf.int[2]

MalelowProb <- Malelow / (1 + Malelow)
MalelowEffect <- MalelowProb * 100

Malehighprob <- Malehigh / (1 + Malehigh)
MalehighEffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1 + MaleEstimate)
MaleEffect <-Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Males is `r MaleEffect`. The confidence interval is `r MalelowEffect` to `r MalehighEffect`

# Females

# Drug Users

# LGBTQ

# Conclusions and recommendations

