---
title: "Covid19 vaccine study Fall2021"
author: "Paulo Poblete"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```

## Barplot 
```{r}
barchartGC(~treatment + infected ,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")
```

As per the plot above, it shows us that those who took the placebo got more sick than the males who took the Covid91. We can surmise that the "sick" portion of the placebo group is almost double compared to the drug group. This shows that the drug is effective for the Males group. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
As per the table above, 33.11 of the sick males had the treatment, and 66.89 of those males who got sick had the placebo. There is a 2:1 difference. This lets us know that the treatment is indeed working. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

As per the chi sq test above, the p value is very tiny. This suggest that we need to reject the null hypothesis and that there is a difference between the males who took the drug in comparison to those who took the placebo. 

## calculating effectiveness

```{r}
MaleEstimate <- (Malesmod$estimate + 1)*100
Malelow <- Malesmod$conf.int[1]
Malehigh <-Malesmod$conf.int[2]

MalelowProb <- Malelow / (1 + Malelow)
MalelowEffect <- MalelowProb * 100

Malehighprob <- Malehigh / (1 + Malehigh)
MalehighEffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1 + MaleEstimate)
MaleEffect <-Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Males is `r MaleEffect`. The confidence interval is `r MalelowEffect` to `r MalehighEffect`

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```

## Barplot 
```{r}
barchartGC(~infected + infected ,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

As per the plot above, the vaccine is not effective for "Females". The Drug portion indicates that out of those who are sick had taken the vaccine. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```
As per the table above, it shows that the females who got sick, 58% of them had gotten the vaccine, in comparison with 42% of the sick that took the placebo. Upon reviewing the information, we can clearly see that the vaccine does not seem to be effective for Females. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

Since our p-value is tiny, we have to reject the null hypothesis, stating that there is a difference between our data. It looks like the data is different in a negative way, the drug is not effective in comparison with the placebo. 

## calculating effectiveness

```{r}
FemaleEstimate <-(Femalesmod$estimate+1) * 100
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <-Femalesmod$conf.int[2]

FemalelowProb <- Femalelow / (1 + Femalelow)
FemalelowEffect <- FemalelowProb * 100

Femalehighprob <- Femalehigh / (1 + Femalehigh)
FemalehighEffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1 + FemaleEstimate)
FemaleEffect <-Femaleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Females is `r FemaleEffect`. The confidence interval is `r FemalelowEffect` to `r FemalehighEffect`

# Drug Users
```{r}
DrugUser <- filter(Covid91, DrugUser == "yes")
```

## Barplot
```{r}
barchartGC(~infected + infected ,data= DrugUser)
barchartGC(~infected + treatment,data= DrugUser, type="percent")
```

As per the barplot above, it shows that the drug is effective on the Drug users.Out of those who took the drug, the placebo group who took had way more infection reaching more than 60% while those who took the drug only made it to about 38%. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

As per the chart above, we can get a more accurate result. It shows that 33% of those who took the drug got infected while 67% of those who took the placebo get infected. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUsersmod <- fisher.test(table1)
```

As per the chi sq test, the p value is (6e-04) we are going to have to reject the null hypothesis. There is a difference between the drug users who took the drug and infected, and those who took the placebo. A positive kind of difference but a difference non the less. 

## calculating effectiveness

```{r}
DrugUserEstimate <-(DrugUsersmod$estimate +1)*100
DrugUserlow <- DrugUsersmod$conf.int[1]
DrugUserhigh <-DrugUsersmod$conf.int[2]

DrugUserlowProb <- DrugUserlow / (1 + DrugUserlow)
DrugUserlowEffect <- DrugUserlowProb * 100

DrugUserhighprob <- DrugUserhigh / (1 + DrugUserhigh)
DrugUserhighEffect <- DrugUserhighprob * 100

DrugUserprob <- DrugUserEstimate / (1 + DrugUserEstimate)
DrugUserEffect <-DrugUserprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Drug Users is `r DrugUserEffect`. The confidence interval is `r DrugUserlowEffect` to `r DrugUserhighEffect`

# LGBTQ

```{r}
LGBTQ <- filter(Covid91, LGBTQ == "gay")
```

## Barplot 

```{r}
barchartGC(~infected + infected ,data= LGBTQ)
barchartGC(~infected + treatment,data= LGBTQ, type="percent")
```

As per the Bar plot above, the LGBTQ group does not gain anything from taking the drug. In the sick portion of the graph, more than 50% got infected even when they took the drug in comparison with the placebo group. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

This chart demonstrates to us that the LGBTQ should avoid taking this drug since, 71 % of those who got sick took the drug in comparison with the 29% of the sick people who handt.  

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQsmod <- fisher.test(table1)
```

As per the chi sq test, the p-value is tiny. This shows that we have to reject the null hypothesis. Also it shows that there is a significant contrast between the LGBQT who took the drug. In this particular case the difference is negative, there were more people who got sick while they took the drug.

## calculating effectiveness

```{r}
LGBTQEstimate <-(LGBTQsmod$estimate+1)*100
LGBTQlow <- LGBTQsmod$conf.int[1]
LGBTQhigh <-LGBTQsmod$conf.int[2]

LGBTQlowProb <- LGBTQlow / (1 + LGBTQlow)
LGBTQlowEffect <- LGBTQlowProb * 100

LGBTQhighprob <- LGBTQhigh / (1 + LGBTQhigh)
LGBTQhighEffect <- LGBTQhighprob * 100

LGBTQprob <- LGBTQEstimate / (1 + LGBTQEstimate)
LGBTQEffect <-LGBTQprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the LGBTQ is `r LGBTQEffect`. The confidence interval is `r LGBTQlowEffect` to `r LGBTQhighEffect`

# Conclusions and recommendations
In conclusion, we had to go do a statistical report on each of the group that 
consist of, Males, Females, Drug users and LGBTQ group. 

As for the Males we found out that the drug was effective for Males group. 
It yields the result of 33% of those who took the drug got sick in comparison with 
the 66.89% of those who had the placebo got sick. the p-value is tiny so we had to 
reject the null hypothesis. While the effectiveness of the Covid91 vaccines for the
Males is `r MaleEffect`. The confidence interval is `r MalelowEffect` to `r 
MalehighEffect`

For the Females group we found out that the drug was ineffective. The numerical 
results showed us that 58% of the women got sick took the drug, in comparison with 
42% of the sick that took the placebo. The p-value is small.We have to reject the 
null hypothesis, stating that there is a difference between our data. It looks like
the data is different in a negative way, the drug is not effective in comparison 
with the placebo. While the effectiveness of the Covid91 vaccines for the Females 
is `r FemaleEffect`. The confidence interval is `r FemalelowEffect` to `r 
FemalehighEffect`

For the drug user group we found out that the drug was pretty effective. The 
numerical results show is that only 33% of those who took the drug got infected in 
comparison to the 67% who those who took the placebo got infected. The p-value is 
0.003 so we have to reject it.  There is a difference between the drug users who 
took the drug and infected, and those who took the placebo. While the effectiveness
of the Covid91 vaccines for the Drug Users is `r DrugUserEffect`. 
The confidence interval is `r DrugUserlowEffect` to `r DrugUserhighEffect

For the LGBTQ group we found out that the drug was ineffective for them. the 
numerical results show 71 % of those who got sick took the drug in comparison with 
the 29% of the sick people who taken the placebo. The p-value is pretty small so we
have to reject the null hypothesis. his shows that we have to reject the null 
hypothesis. Also it shows that there is a significant contrast between the LGBQT 
who took the drug. In this particular case the difference is negative, there were 
more people who got sick while they took the drug. While the effectiveness of the 
Covid91 vaccines for the LGBTQ is `r LGBTQEffect`. The confidence interval is `r 
LGBTQlowEffect` to `r LGBTQhighEffect`
