---
title: "Covid19 vaccine study Fall2021"
author: "Paulo Poblete"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```

## Barplot 
```{r}
barchartGC(~treatment + infected ,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")
```

As per the plot above, it shows us that those who took the placebo got more sick than the males who took the Covid91. We can surmise that the "sick" portion of the placebo group is almost double compared to the drug group. This shows that the drug is effective for the Males group. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
As per the table above, 33.11 of the sick males had the treatment, and 66.89 of those males who got sick had the placebo. There is a 2:1 difference. This lets us know that the treatment is indeed working. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

As per the chi sq test above, the p value is very tiny. This suggest that we need to reject the null hypothesis and that there is a difference between the males who took the drug in comparison to those who took the placebo. 

## calculating effectiveness

```{r}
MaleEstimate <-Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <-Malesmod$conf.int[2]

MalelowProb <- Malelow / (1 + Malelow)
MalelowEffect <- MalelowProb * 100

Malehighprob <- Malehigh / (1 + Malehigh)
MalehighEffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1 + MaleEstimate)
MaleEffect <-Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Males is `r MaleEffect`. The confidence interval is `r MalelowEffect` to `r MalehighEffect`

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```

## Barplot 
```{r}
barchartGC(~infected + infected ,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

As per the plot above, the vaccine is not effective for "Females". The Drug portion indicates that out of those who are sick had taken the vaccine. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```
As per the table above, it shows that the females who got sick, 58% of them had gotten the vaccine, in comparison with 42% of the sick that took the placebo. Upon reviewing the information, we can clearly see that the vaccine does not seem to be effective for Females. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

Since our p-value is tiny, we have to reject the null hypothesis, stating that there is a difference between our data. It looks like the data is different in a negative way, the drug is not effective in comparison with the placebo. 

## calculating effectiveness

```{r}
FemaleEstimate <-Femalesmod$estimate
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <-Femalesmod$conf.int[2]

FemalelowProb <- Femalelow / (1 + Femalelow)
FemalelowEffect <- FemalelowProb * 100

Femalehighprob <- Femalehigh / (1 + Femalehigh)
FemalehighEffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1 + FemaleEstimate)
FemaleEffect <-Femaleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccines for the Females is `r FemaleEffect`. The confidence interval is `r FemalelowEffect` to `r FemalehighEffect`

# Drug Users

# LGBTQ

# Conclusions and recommendations

